\chapter{Preliminaries}

\section{Elements of Operator Algebras}

A \(C^\ast\)-algebra \(A\) is a Banach algebra over the field \(\mathbb{C}\) together with a map \(\ast: A \to A\), \(a \mapsto a^\ast\) which following properties:
\begin{enumerate}
    \item \(a^{\ast\ast}=a\) for all \(a \in A\),
    \item \((a+b)^\ast=a^\ast+b^\ast\) for all \(a,b \in A\),
    \item \((ab)^\ast = b^\ast a^\ast\) for all \(a,b \in A\),
    \item \((\lambda a)^\ast = \bar{\lambda}a^\ast\) for all \(\lambda \in \mathbb{C}\) and \(a \in A\),
    \item \(\left \lVert aa^\ast \right \rVert = \left \lVert a \right \rVert \left \lVert a^\ast \right \rVert\) for all \(a \in A\).
\end{enumerate}

A bounded linear map \(\pi:A \to B\) between \(C^\ast\)-algebras \(A\) and \(B\) is called a \(\ast\)-homomorphism if
\begin{enumerate}
    \item \(\pi(ab)=\pi(a)\pi(b)\)
    \item \(\pi(a^\ast) = \pi(a)^\ast\)
\end{enumerate}
for all \(a,b \in A\).

Let \(A\) be a \(C^\ast\)-algebra. A \emph{state} in \(C^\ast\)-algebra is a positive functional \(\rho\) which is \(\left \lVert \rho \right \rVert=1\). We shall denote the set of all states on \(A\) by \(S(A)\) which is called the \emph{state space} of A.

A \(\ast\)-representation of a \(C^\ast\)-algebra \(A\) on a Hilbert space \(H\) is a ring homomorphism \(\pi: A \to B(H)=\{f \mid f \mbox{ is a bounded linear operator on }H\}\) such that
\begin{enumerate}
    \item \(\pi(a^\ast)=\pi(a)^\ast\) for all \(a \in A\),
    \item \(\pi\) is nondegenerate.
\end{enumerate}

\begin{theorem}[GNS representation]
    Given a state \(\rho\) of \(A\), there is a \(\ast\)-representation \(\pi\) of \(A\) acting on a Hilbert space \(H\) with unit cylic vector \(\psi\) such that
    \[\rho(a)=\left \langle \pi(a) \psi,\psi  \right \rangle\]
    for every \(a\) in \(A\). Moreover, the representation is unique up to unitary equivalence.
\end{theorem}
\begin{proof}
    \cite[Theorem 4.5.2]{MR1468229}
\end{proof}

Let \(X\) be an inner product space. A \emph{spilitting subspace} of \(X\) is the subspace \(M\) of \(X\) which is the following property holds,
\[M+M^\perp=S.\]
Then any vector \(x \in X\) can be uniquely expressed in the form
\[x=x_M+x_{M^{\perp}},\]
where \(x_M \in M\) and \(x_{M^\perp} \in M^\perp\). We denote by \(E(X)\) the set of all splitting subspaces.
The map \(P_M:X\to X\) such that \(P_M(x)=x_M\) for all \(x \in X\) is a bounded linear operator with \(\left \lVert P_M \right \rVert=1\) whenever \(M \neq \{0\}\). Moreover, \(P^2_M=P_M\), \(P_M\) is self-adjoint, and \(\mathrm{ran}\, P_M = M\). The operator \(P_M\) is called the \emph{orthoprojector} or \emph{projection} from \(X\) onto \(M\). It is clear that \(I-P_M\) is an orthoprojector onto \(M^\perp\).
\begin{proposition}
    Let \(P\) is an idempotent linear operator on \(X\) which is self-adjoint. If \(M=\{x\in X \mid Px=x\}\), then \(M \in E(S)\) and \(P_M = P\).
    \begin{proof}
        We have
        \[\left \langle x,z-Pz \right \rangle=\left \langle x,z \right \rangle-\left \langle x,Pz \right \rangle=\left \langle x-Px,z \right \rangle=0\]
        for all \(x \in M\) and for all \(z \in X\). Therefore, \(z-Pz \in M^\perp\), and \[z=Pz+(I-P)z\]
    \end{proof}
\end{proposition}
\begin{corollary}
    If \(P\) is the orthoprojector onto a splitting subspaces \(M\) of \(X\), then \(I_P\) is the orth projector onto \(M^\perp\) and \(M^\perp=\{x\in X \mid Px=0\}\).
\end{corollary}
\begin{proposition}
    Let \(\{e_1,\dots,e_n\}\) is a finite orthonormal system in \(X\). Then \(M=\{\mathrm{sp}(e_1,\dots,e_n) \in E(S)\}\) and
    \[P_M(x) = \sum_{i=1}^{n} (x,e_i)e_i\]
    for all \(x \in X\).
\end{proposition}
\begin{lemma}
A splitting subspace \(M\) of \(X\) is invariant under an operator \(T\) on \(X\) if and only if \(AP_M=P_MTP_M\).
\end{lemma}
\begin{lemma}
Let \(M\) and \(N\) be two splitting subspaces. Then the followings are equivalent.
\begin{enumerate}
    \item \(M \perp N\).
    \item \(P_M P_N = O\)
\end{enumerate}
\begin{proof}
    If \(M\perp N\), then \(N \subset M^\perp\). We have \(P_Nx \in N\) for all \(x \in X\), so that \(P_MP_N=0\) for all \(x \in X\). If \(P_M P_N=O\), conversely, then \(P_Mx=P_M P_N x =0\) for all \(x \in N\). Therefore \(N \subset M^\perp\), so that \(M\perp N\).
\end{proof}
\end{lemma}
We say that two orthoporjectors \(P\) and \(Q\) on \(X\) are \emph{orthogonal} if \(PQ=O\), and we write \(P\perp Q\).
\begin{lemma}
    Let \(M\) and \(N\) be two splitting subspaces of \(X\). The following statements are equivalent:
    \begin{enumerate}
        \item \(P_M \le P_N\).
        \item \(\left \lVert P_M \right \rVert \le \left \lVert P_Nx \right \rVert\) for all \(x \in X\).
        \item \(M \subset N\).
        \item \(P_M P_N = P_M\).
        \item \(P_N P_M = P_M\).
    \end{enumerate}
\end{lemma}

\section{Quantum Logic}

Let \(L\) be a poset. For a sequence \((a_i)_{i\in I}\) in \(L\), the \emph{join} is defined by \[\bigvee_{i \in I} a_i = \sup \{a_i \mid i \in I\},\]
and the \emph{meet} is defined by
    \[\bigwedge_{i \in I} a_i = \inf \{a_i \mid i \in I\}.\]
    \begin{definition}
        Let \(L\) be a poset. Then
        \begin{itemize}
            \item \(L\) is a \emph{lattice} if \(a\vee b\) and \(a \wedge b\) exist in L for any \(a,b \in L\).
            \item \(L\) is a \emph{\(\sigma\)-lattice} if \(\bigvee_{i \in I} a_i\) and \(\bigwedge_{i\in I} a_i\) exist in \(L\) for any sequence \((a_i)_{i\in I}\) with a countable index set \(I\).
            \item \(L\) is a \emph{complete lattice} if \(\bigvee_{i \in I} a_i\) and \(\bigwedge_{i \in I} a_i\) exist in \(L\) for any sequence \((a_i)_{i\in I}\) with an arbitrary index set \(I\).
        \end{itemize}
    \end{definition}
\begin{definition}
    A poset \(L\) is said to be \emph{bounded} if there exist elements \(0\) and \(1\) in \(L\),
    \[0\le x \le 1\]
    for all \(x \in L\).
\end{definition}
A unary operation \(\perp: L \to L, a \mapsto a^\perp\) is said to be an \emph{orthocomplementation} on a bounded poset \(L\) if
\begin{enumerate}
    \item \((a^\perp)^\perp=a\),
    \item if \(a \le b\), then \(b^\perp\le a^\perp\),
    \item \(a\vee a^\perp=1\)
\end{enumerate}
for any \(a,b \in L\). A bounded poset \(L\) with the orthocomplementation \(\perp\) is called \emph{orthocomplemented}.
\begin{definition}
    An orthocomplemented poset \(L\) is said to be an \emph{orthomodular poset} (in short OMP) if
    \begin{enumerate}
        \item \(a \vee b \in L\) if \(a\perp b\),
        \item (orthomodularity) \(b=a\vee(b \wedge a^\perp)\) if \(a \le b\),
    \end{enumerate}
    hold for all \(a,b \in L\).
\end{definition}
We denote an orthomodular lattice by OML. If an OMP \(L\) have the following property
\[\bigvee_{i=1}^\infty a_i \in L\]
for any pairwise orthgonal sequence \((a_i)_{i=1}^{\infty}\),
\(L\) is called a \emph{quantum logic}.
\begin{definition}
    Let \(L_1\) amd \(L_2\) be two OMPs. A mapping \(h:L_1 \to L_2\) is called a \emph{homomorphism} if
    \begin{enumerate}
        \item \(h(1)=1\),
        \item \(h(a)\perp h(b)\) if \(a\perp b\) in \(L_1\),
        \item \(h(a\vee b)=h(a)\vee h(b)\) if \(a\perp b\) in \(L_1\).
    \end{enumerate}
    In addition, we call \(h\) a \(\sigma\)-homomorphism if
    \[h(\bigvee_{i=1}^{\infty} a_i) = \bigvee_{i=1}^{\infty} h(a_i)\]
    holds for any sequence of pairwise orthogonal elements \((a_i)_{i=1}^{\infty}\) of \(L_1\).
\end{definition}
\begin{definition}
    Let \(L\) be an OMP. Consider a map \(\mu:L\to [-\infty,\infty]\) such that
    \begin{gather}
        \mu(0)=0,\\
        \mu(\bigvee_{i \in I} a_i) = \sum_{i \in I} \mu(a_i), \mbox{ for any pairwise orthogonal sequence } (a_i)_{i \in I}.
    \end{gather}
    Then,
    \begin{enumerate}
        \item \(\mu\) is called a \emph{finitely additive measure} or a \emph{charge} if (1.2) holds for any finite set \(I\),
        \item \(\mu\) is called a \emph{signed measure} if (1.2) holds for any countably infinite set \(I\),
        \item \(\mu\) is called a \emph{completely additive signed measure} if (1.2) holds for any arbitrary set \(I\).
    \end{enumerate}
If \(\mu\) is nonnegative and \(\mu(1)=1\), then we call \(\mu\) a \emph{state}.
\end{definition}

\begin{example}
    Let \(H\) be a Hilbert space over a real or complex. The class
    \[L(H)=\{M \mid M \mbox{ is a closed subspace of } H\}\]
    is a quantum logic, where the partial ordering is the inclusion, the meet and join are defined as follows
    \[\bigwedge_{i \in I} M_i, \quad \bigvee_{i \in I} M_i = \mathrm{cl}(\mathrm{span(\bigcup_{i\in I} M_i)}),\]
    and the orthocomplementation \(\perp:M \mapsto M^\perp\).
\end{example}

\begin{example}
    Let \(H\) be a Hilbert space over a real or complex. The class
    \[\mathcal{P}(H)=\{P \mid P \mbox{ is a orthoprojector on } H\}\]
    is a quantum logic where \(P\le Q\) if and only if \[\left \langle Px,x \right \rangle \le \left \langle Qx,x \right \rangle\]
    for all \(x \in H\) and \(P^\perp=I-P\).
\end{example}